---
title: "HW 6"
output: github_document
date: "2025-11-17"
author: "Julia Gray"
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Get data:

```{r}
homicide_df = read.csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") |> 
  janitor::clean_names() |> 
  mutate(
    reported_date = as.Date(as.character(reported_date), format = "%y%m%d"),
    city_state = paste(city, state, sep=", "),
    solved = as.numeric(disposition %in% c("Closed by arrest")),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")
  ) |> 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    !victim_age %in% c(NA),    #remove age is NA
    victim_race %in% c("White", "Black")
  )

baltimore_df = homicide_df |> 
  filter(
    city_state == "Baltimore, MD"
  ) |> 
  select(solved, victim_age, victim_race, victim_sex)
```

Fit model for Baltimore:

```{r}
or_balt = 
  baltimore_df |> 
  glm(solved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) 

or_balt |> 
  broom::tidy(conf.int = TRUE, exponentiate = TRUE) |> 
  select(term, estimate, conf.low, conf.high, p.value) |> 
  knitr::kable(digits = 3)
```

Fit model for all cities:

```{r}
fit_or = function(input_city){
  
  fit = 
    homicide_df |>
    filter(city_state == input_city) |>
    glm(formula = solved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) |>
    broom::tidy(conf.int = TRUE, exponentiate = TRUE) |>
    filter(term == "victim_sexMale") |>
    select(estimate, conf.low, conf.high, p.value)

  fit
}

ny_test = fit_or('New York, NY')

or_all_cities = 
  tibble(
    city_s = unique(pull(homicide_df, city_state))
    ) |> 
  mutate(fit_estimates = map(city_s, fit_or)) |> 
  unnest(fit_estimates)
```

Plot the results

```{r}
or_all_cities |> 
  mutate(city_s = fct_reorder(city_s, estimate)) |> 
  ggplot(aes(x = estimate, y = city_s)) + 
  geom_point() + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 1, color = "red") +
  xlim(0, 4) +
  ylab("City, State") +
  xlab("OR of homicide being solved (male vs. female)")
```

An odds ratio of 1 (shown on the plot as a red line), means there is no difference in the odss of the homicide being solved between males and females. An odds ratio greater than 1 means that homicides where the victim is male have higher odds of being solved and vice versa. The confidence interval is shown by the error bars and where it crosses 1 the results are not significant. 

## Problem 2

